Refactoring is a fundamental activity in software engineering to reorganize code to improve its structure, e.g., to simplify maintenance, while preserving its observable behavior of the program.
A refactoring can be defined as a pattern it matches on, and a subsequent program transformation on the matched part.
To ensure that the transformed program indeed has the same observable behavior one can either compare the transformed program with the 
original or %through, e.g., regression testing~\cite{regression} or relational verification~\cite{relate}, or 
reason about the program transformation itself.

Relational verification through dynamic logic is a promising approach to verification of refactoring patterns in object oriented programs.
Recent advances from symbolic to Abstract Execution (AE)~\cite{steinhoefel:ae} have enabled reasoning about incomplete/abstract versions of such programs.
This has proven fruitful in the exploration of correctness of refactorings primarily related to code blocks in Java:
%In this paper we explore further types of equivalent transformations and refactorings and discuss the challenges that still need to be overcome for full round-trip correctness of refactorings in object-oriented languages.
%
AE introduces abstract statements (and expressions), which act as named and specified placeholders for statement-sequences of the host language.
A refactoring proof is a relational verification proof that compares two programs which can have abstract program elements.
Consider relating
\begin{center}
\jmcode{if (E$_{\mathit{boolean}}$) \{S$_1$;\} else \{S$_2$;\} return;}\\[0.3em]
and\\[0.3em]
\jmcode{if (!E$_{\mathit{boolean}}$) \{S$_2$; return;\} S$_1$; return;}
\end{center}
where \jmcode{E$_{\mathit{boolean}}$} is an arbitrary boolean expression, \jmcode{S$_1$} and \jmcode{S$_2$} correspond to arbitrary statements.
One of the programs represents the schema of the code before the refactoring, and the other one the code afterwards, with respect to some relational post-condition that defines the notion of \emph{program equivalence}. This reasoning about abstract programs is exactly the aforementioned reasoning about the program transformation behind the refactoring.

Despite the promising results due to AE, it has only been applied to \emph{statement-level} refactorings that
change the body of a single method.
However, refactorings are not limited to single methods or statement-blocks, but may also restructure classes and data structures \cite{fowler:refactoring}.
In this paper we investigate the role of AE for verification of refactorings beyond statement-level.


\paragraph{Challenges.}
More complex refactorings require more elaborate specification and verification techniques for relational verification. 
The reason is that the programs surrounding the abstract statements, as well as the notion of equivalence, become more involved.
This holds for both structure and behaviour.
Following the general approach of AE, \Refinity{} \cite{steinhoefel:ae}, the tool supporting verification on top of the KeY system, an automated theorem prover for Java~\cite{DBLP:conf/aplas/Steinhofel20},
has been primarily designed to verify the correctness of refactorings that are based on moving code within a method, or on extracting some statements into their own method (i.e., the \rname{Extract Method} refactoring~\cite[p. 106]{fowler:refactoring2nd}).

To investigate the use of AE for refactorings on the class-level, we investigate the \rname{Hide Delegate} refactoring~\cite[p. 189]{fowler:refactoring2nd} that moves code between classes and show how it can be encoded in \Refinity{}.
As for specification, we discuss the interpretation of \textit{equivalence} from the perspective of the user and how to encode this --- for example, if the surrounding programs throw exceptions, under which conditions are the exceptions considered equivalent?
The equivalence interpretation goes beyond exceptions, but touches on a fundamental problem:
there are several possible choices for when newly created objects (and exceptions, which are objects in Java) are considered to be equal in relational verification using dynamic logic.
This was first investigated by Beckert et al.~\cite{DBLP:conf/lopstr/BeckertBKSSU13} and we discuss alternatives here.


For exceptions and object creation, we describe several possibilities when newly created objects (resp.\ thrown exceptions) are considered equal
and how this information can be used in dynamic logic proofs. The different possibilities are implemented as 
%\vsnote*{Unclear, at least that early in the paper. Why ``different''? Also see my later point about why using the new rule(s) should be optional.}{different rule sets} 
\EK{multiple (sets of) rules, from which the developer chooses the one corresponding to his assumptions on object allocations -- this choice
does not have to be encoded explicitly in the relational post-condition.}
This reduces the size of the required specification, which is an advantage, since it is a notorious bottleneck in formal verification~\cite{DBLP:journals/corr/abs-1211-6186,DBLP:series/lncs/HahnleH19}.


Furthermore, we discuss the necessary extensions needed to prove equivalent behaviour where one data structure is replaced by another, e.g., an array or any primitive type by a class.
Here, the main challenge lies in the encoding that the structures are used correctly throughout the execution.
This could, for example, be handled by coupled invariants~\cite{DBLP:conf/birthday/BeckertU18}, whose connection to AE is yet unclear. 
Lastly, we discuss the challenge to apply AE to novel specification approaches for \emph{traces} which aim to simplify the specification of temporal properties for expressive properties, but whose use for relational verification is unexplored.

\paragraph{Contributions and Structure.}
Our contributions include an investigation into the necessary side conditions to be able to proof of the correctness of the \rname{Extract Local Variable}- and \rname{Hide Delegate} refactorings, extending the collection of proven refactorings.
The latter is a refactoring beyond code motion within a method and highlights the interaction of AE with general relational verification challenges.
Then, we discuss possible extensions that would be required to address further refactorings with AE.

%This paper is structured as follows. 
We first describe AE and relational verification using the \rname{Extract Local Variable} refactoring in Section~\ref{sec:prelim}, before we show necessary conditions for the \rname{Hide Delegate} refactoring to be correct  and then discuss the challenges for AE in Section~\ref{sec:challenges}.
\vsnote*{@ALL}{Sec.~\ref{sec:future} proposes future directions for improvements in REFINITY and AE which most likely require major development effort.}
In Section~\ref{sec:discussion}, we discuss our results. We discuss the related work in  Section~\ref{sec:related}, and lastly conclude in Section~\ref{sec:conclusion}.



%This paper is structured as follows: we first provide an overview on abstract executions, the main vehicle for correctness- and equivalence proofs, and their use in REFINITY.
%Next, we provide an encoding of the Hide Delegating refactoring for REFINITY that moves away from statement-based refactorings to object-oriented refactorings in Section \ref{sec:hideDelegate}.
%After that in Section \ref{sec:challenges}, we explore (and address, where possible) \vsnote*{Not really limitations...better word?}{limitations} in REFINITY that we have discovered by investigating further refactorings.
%Our main goal is to extend the notion of \textit{equal} programs to \textit{equivalent} programs, e.g. by replacing data types.
%\vsnote{Mention possible application to optimisations already in intro? Probably.}
%We conclude with a discussion in Section \ref{sec:discussion} on the feasibility of necessary developments \vsnote*{esp. if we want to use histories in subsequent proofs.}{and their dependencies}.
% VS: This will then be included in the section above:
% Next we discuss issues related to formally correct and constructive handling of object creation to be able to express useful notions of equivalence in Section \ref{sec:objectcreation}.
% Section \ref{sec:trace} moves away from only heap-related equivalence and discusses tracking equivalent observable behaviour in a relational framework.
% The final challenge are relational invariants that we discuss in Section \ref{sec:relinv} which are required to track equivalent, yet structurally different representation of data types.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
